Concentration Inequalities

### 2016-06-08 TPR/TPR_report.tex
cleaned up structure of causality folder
adds initial writup of contextual-bandit and identifiability problems

\section{Introduction}

@@

\section{Causal Inference}

@@

\section{Unifying the frameworks}

@@

\subsection{Incorporating side information induced by casual graph}

As a first step towards more general problems, we develop and evaluate a bandit algorithm that incorporates the side information induced by a specific causal graph.

Assume we have a known causal model with binary variables $\boldsymbol{X} = \{X_{1}..X_{N}\}$ that independently cause a target variable of interest $Y$, figure \ref{fig:causalStructure}. This is an expansion of the model shown in figure \ref{fig:unify_frameworks}. We can run sequential experiments on the system, where at each timestep $t$ we can either do nothing, $do()$, or select a single variable, $X_i$, on which to intervene, $do(X_{t,i} = J_t)$, and subsequently observe the complete result, $(\boldsymbol{X}_{t},Y_{t})$.

As an example, consider a farmer wishing to optimize the yield of her crop. She can invest in a green house to control temperature, a watering system to control soil moisture, fertilizers to set soil nutrients, etc. We assume only a single intervention is feasible due to cost and that each of these variables are independent of one-another (this may not always be the case - temperature could be related to rainfall for example). After having selected which variable to control, she plants her crops and observes the values of the remaining input variables and the yield. This repeats across many growing seasons, and the goal is to maximize the total cumulative yield.

\begin{figure}[h]
\centering
\caption{Assumed Causal Structure}
\label{fig:causalStructure}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
  thick,main node/.style={observed}, hidden/.style={empty},tg/.style={target}]

 %nodes
\node[main node](1){$X_{1}$};
\node[main node, right=of 1](2){$X_{2}$};
\node[hidden, right=of 2](3){$...$};
\node[main node, right=of 3](4){$X_{N}$};
\node[tg, below right=of 2](5){Y};
 \path[every node/.style={font=\sffamily\small}]
    (1) edge (5)
    	(2) edge (5)
    (4) edge (5);

\end{tikzpicture}
\end{figure}

Let $\boldsymbol{q} \in [0,1]^N$ be a fixed vector where $q_i = P(X_i = 1)$. In each time-step $t$ upto a known horizon $T$:

\begin{enumerate}
\item The learner chooses an $I_t \in \set{0,\ldots, N}$ and $J_t \in \set{0,1}$, where $I_t = 0$ represents $do()$.

\item Variables not set by the learner take their values probabilistically according to $\boldsymbol{q}$ 
\eq{
X_{t,i} = \begin{cases}
\sim \Ber(q_i) &\text{if } i \neq I_t \\
J_t & \text{otherwise}\,.
\end{cases}
}
\item The learner observes $(\boldsymbol{X}_{t},Y_{t})$ and receives reward $Y_t \sim \Ber(r(\boldsymbol{X}_{t}))$ where $r:\set{0,1}^N \to [0,1]$ is unknown and arbitrary. 
\end{enumerate}

The problem could be treated as a standard bandit problem with $K = 2N + 1$ arms,

\eq {
\{do(), do(X_1=0), do(X_1 = 1)\;\; ... \;\;do(X_N = 1)\}
}

yielding a regret $\bigo{\sqrt{TN}}$. However, this doesn't leverage the side information induced by the causal structure:

\eqn {
\label{eq:observe}
P(Y_t|do(X_{i,t} = j)) &= P(Y_t|do(),X_{i,t} = j) \\
\label{eq:estimation_transfer}
&= P(Y_t|do(X_{l,t}=0),X_{i,t}=j)(1-q_l)+P(Y_t|do(X_{l,t}=1),X_{i,t}=j)q_l 
}

The hardness of this problem depends strongly on the vector $\boldsymbol{q}$. This is intuitive from a consideration of the extreme cases. If $\boldsymbol{q} = \boldsymbol{0}$ (or $\boldsymbol{1}$), then half of the actions we wish to explore will only occur if we explicitly select them. For these actions we get no side information by playing other actions. So at best we could expect a factor of two improvement over simply running a standard bandit algorithm. If $\boldsymbol{q} = \boldsymbol{\frac{1}{2}}$, the observe action, $do()$, effectively becomes a revealing action and we would anticipate obtaining a regret with minimal dependence on the number of variables $N$.

We consider a simple explore-exploit based algorithm, that will explore for $h$ timesteps, sampling actions in a way that depends on $\boldsymbol{q}$. We then select the arm with the highest expected reward for the remaining $T-h$ time steps. During the exploration phase, we will balance purely observing, $I_t = 0$, which takes advantage of equation \ref{eq:observe} to reveal the reward for frequently occurring arms, with explicit sampling of infrequently occurring arms.

Without loss of generality, we can assume $q_i \in [0,\frac{1}{2}]$ and $q_1 \leq q_2 ... \leq q_N$. Let:

\eqn {
m \in [2,N] = \set{m:q_m > \frac{1}{m}}
}

$m$ is a measure of the number of infrequently occurring arms. If the problem is completely balanced $q_1...q_N = \frac{1}{2}$ then $m = 2$. If the problem is completely unbalanced, $q_1...q_N = 0$ then $m=N$.  Let $q_{ij} = \P{X_i = j}$ and $\mu_{ij} = P(Y|do(X_i = j))$.

Suppose we select $I_t = 0$ for the first $h/2$ timesteps. This is at worst half the optimal. We use these observations to estimate the rewards for all the frequently occurring arms, $\set{do(X_i = 0) \; \forall i}$, and $\set{X_i = 1:i \geq m}$

\eqn {
\hat{\mu}_{ij} = \frac{\sum_{t=1}^{h/2}\ind{Y_t=1,X_{t,i}=j}}{\frac{h}{2}q_{ij}}
}

We then explicitly play each of the $m$ infrequently occurring arms $\frac{h}{2m}$ times and estimate their rewards as:

\eqn {
\hat{\mu}_{ij} = \frac{2m}{h}\sum_{t = 1}^{h/2m}\ind{Y_t = 1|X_i=j}
}

We then simply play the arm with the highest estimated reward for the remaining $T-h$ timesteps. We can show that we can select an $h$ (based on $T$, $N$ and $m$) that leads to worst case regret $R_T = \bigotilde{T^{2/3}m^{1/3}}$. This result can still be obtained in the case where $\boldsymbol{q}$ is unknown and must be learned during the exploration time. Comparing this to the standard bandit regret, $R_T = \bigo{\sqrt{TK}}$, we expect the causal algorithm to do better if $m < \frac{K^{3/2}}{\sqrt{T}}$. We can also show that the simple regret $R_s \in \bigotilde{\sqrt{m/T}}$. If $m << K$, this represents a significant improvement over standard bandit setting. \footnote{ $\bigotilde{.}$ and $\bigthetatilde{.}$ suppress polylogarithmic factors}

In this algorithm, we have only used the side information provided by the $do()$ action about other actions. Since the $do()$ action fully reveals the value of alternate actions we could have incorporated this information via the graph feedback model \cite{Mannor2011}, where at each timestep the feedback graph $G_t$ is selected stochastically, dependent on $\boldsymbol{q}$, and revealed after an action has been chosen. The feedback graph is distinct from the causal graph. A link $A \rightarrow B$ in $G_t$ indicates that selecting the action $A$ reveals the reward for action $B$. For this specific problem, $G_t$ will always be a star graph with the action $do()$ connected to half the remaining actions. The Exp3-IX algorithm \cite{Kocak2014} was developed for the adversarial version of this problem and has regret $\bigo{\sqrt{\bar{\alpha}T}}$, where $\bar{\alpha}$ is the average independence number of $G_t$. In our case $\bar{\alpha} = \frac{N}{2}$ so we again obtain the regret of the standard bandit algorithm. The issue here is that a malicious adversary can select the same graph each time, such that the rewards for half the arms are never revealed by the informative action. This is equivalent to a, nominally, stochastic selection of feedback graph where $\boldsymbol{q} = \boldsymbol{0}$

\cite{Lelarge2012} consider a stochastic version of the graph feedback problem, but with a fixed graph available to the algorithm before it must select an action. In addition, their algorithm is not optimal for all graph structures and fails, in particular, to provide improvements for star like graphs as in our case. \cite{Buccapatnam2014} improve the dependence of the algorithm on the graph structure but still assume the graph is fixed and available to the algorithm before the action is selected.

More generally, assuming causal structure creates more complex types of side information, such as that shown in equation \ref{eq:estimation_transfer}. In this case, selecting one action does not fully reveal an alternate action but provides some information towards an estimate. The quality of the estimate notably depends not only on the number of times that action was selected. For example, to get a good estimate for $X_1 = 1$ by intervening on $X_2$ requires us to sample both $X_2=0$ and $X_2=1$, in proportions dependent on $q_2$. This more complex side information does not fit within the graph feedback framework.

Partial monitoring is a very general framework for for decoupling the feedback from the action and reward. It can be used to classify problems into one of four categories, trivial with no regret, easy with $R_T = \bigthetatilde{\sqrt{T}}$ , hard with $R_T = \bigtheta{T^{2/3}}$ and hopeless with $R_T = \bigomega{T}$ \cite{Bartok2014}. Partial monitoring algorithms yield results that are optimal with respect the the horizon $T$ but not other parameters, such as $K$, which is the key focus of incorporating causal structure.

\subsection{Open questions}

There are many open questions still to be addressed. For the specific setting described above I need to derive lower bounds for the worst case regret so as to determine how close the current approach is to optimal.

It is clearly desirable to obtain an algorithm that could be generalized to an arbitrary causal graph. This will likely require the development of a bandit algorithm that can leverage more complex forms of side information like that in equation \ref{eq:estimation_transfer}, in addition to an efficient algorithm for eliminating actions that need not be considered at all.

Another open class of problem to consider are cases where the causal structure is not (fully) known and also needs to be learnt within the online setting.

\section{Paid work statement}
I am not currently undertaking or planning to undertake any paid work.

\bibliographystyle{unsrt}% Select the citation style e.g. ieeetr
\bibliography{../library}% write the directory to the .bib file
\end{document}

>>>>

\section{Problem Summary}

@@

\part{Literature review}
\section{Causal Inference}

@@

\part{Approach and expected outcomes}

The goal of my research is to explore a number of questions around learning to make good decisions.

\begin{itemize}
\item Develop a framework to unify concepts causal inference concepts with reinforcement learning, starting with multi-armed bandits.(DONE - see section \ref{sec:unificaition})
\item Demonstrate combining concepts from causal inference with multi-armed bandit algorithms can improve on state of the art performance in sequential decision making (DONE - paper submitted for NIPS 2016)
\item Causal privacy paper

\item something about identifiability
\end{itemize}

\section{Unifying the frameworks}
\label{sec:unificaition}

@@

%\subsection{Incorporating side information induced by casual graph}
%
%As a first step towards more general problems, we develop and evaluate a bandit algorithm that incorporates the side information induced by a specific causal graph. 
%
%Assume we have a known causal model with binary variables $\boldsymbol{X} = \{X_{1}..X_{N}\}$ that independently cause a target variable of interest $Y$, figure \ref{fig:causalStructure}. This is an expansion of the model shown in figure \ref{fig:unify_frameworks}. We can run sequential experiments on the system, where at each timestep $t$ we can either do nothing, $do()$, or select a single variable, $X_i$, on which to intervene, $do(X_{t,i} = J_t)$, and subsequently observe the complete result, $(\boldsymbol{X}_{t},Y_{t})$. 
%
%As an example, consider a farmer wishing to optimize the yield of her crop. She can invest in a green house to control temperature, a watering system to control soil moisture, fertilizers to set soil nutrients, etc. We assume only a single intervention is feasible due to cost and that each of these variables are independent of one-another (this may not always be the case - temperature could be related to rainfall for example). After having selected which variable to control, she plants her crops and observes the values of the remaining input variables and the yield. This repeats across many growing seasons, and the goal is to maximize the total cumulative yield.
%
%\begin{figure}[h]
%\centering
%\caption{Assumed Causal Structure}
%\label{fig:causalStructure}
%\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
%  thick,main node/.style={observed}, hidden/.style={empty},tg/.style={target}]
%
% %nodes
%\node[main node](1){$X_{1}$};
%\node[main node, right=of 1](2){$X_{2}$};
%\node[hidden, right=of 2](3){$...$};
%\node[main node, right=of 3](4){$X_{N}$};
%\node[tg, below right=of 2](5){Y};
% \path[every node/.style={font=\sffamily\small}]
%    (1) edge (5)
%    	(2) edge (5)
%    (4) edge (5);
%	
%\end{tikzpicture}
%\end{figure}
%
%
%Let $\boldsymbol{q} \in [0,1]^N$ be a fixed vector where $q_i = P(X_i = 1)$. In each time-step $t$ upto a known horizon $T$:
% 
%\begin{enumerate}
%\item The learner chooses an $I_t \in \set{0,\ldots, N}$ and $J_t \in \set{0,1}$, where $I_t = 0$ represents $do()$.
%
%\item Variables not set by the learner take their values probabilistically according to $\boldsymbol{q}$ 
%\eq{
%X_{t,i} = \begin{cases}
%\sim \Ber(q_i) &\text{if } i \neq I_t \\
%J_t & \text{otherwise}\,.
%\end{cases}
%}
%\item The learner observes $(\boldsymbol{X}_{t},Y_{t})$ and receives reward $Y_t \sim \Ber(r(\boldsymbol{X}_{t}))$ where $r:\set{0,1}^N \to [0,1]$ is unknown and arbitrary. 
%\end{enumerate}
%
%The problem could be treated as a standard bandit problem with $K = 2N + 1$ arms, 
%
%\eq {
%\{do(), do(X_1=0), do(X_1 = 1)\;\; ... \;\;do(X_N = 1)\}
%}
%
%yielding a regret $\bigo{\sqrt{TN}}$. However, this doesn't leverage the side information induced by the causal structure: 
%
%\eqn {
%\label{eq:observe}
%P(Y_t|do(X_{i,t} = j)) &= P(Y_t|do(),X_{i,t} = j) \\
%\label{eq:estimation_transfer}
%&= P(Y_t|do(X_{l,t}=0),X_{i,t}=j)(1-q_l)+P(Y_t|do(X_{l,t}=1),X_{i,t}=j)q_l 
%}
%
%The hardness of this problem depends strongly on the vector $\boldsymbol{q}$. This is intuitive from a consideration of the extreme cases. If $\boldsymbol{q} = \boldsymbol{0}$ (or $\boldsymbol{1}$), then half of the actions we wish to explore will only occur if we explicitly select them. For these actions we get no side information by playing other actions. So at best we could expect a factor of two improvement over simply running a standard bandit algorithm. If $\boldsymbol{q} = \boldsymbol{\frac{1}{2}}$, the observe action, $do()$, effectively becomes a revealing action and we would anticipate obtaining a regret with minimal dependence on the number of variables $N$. 
%
%We consider a simple explore-exploit based algorithm, that will explore for $h$ timesteps, sampling actions in a way that depends on $\boldsymbol{q}$. We then select the arm with the highest expected reward for the remaining $T-h$ time steps. During the exploration phase, we will balance purely observing, $I_t = 0$, which takes advantage of equation \ref{eq:observe} to reveal the reward for frequently occurring arms, with explicit sampling of infrequently occurring arms. 
%
%Without loss of generality, we can assume $q_i \in [0,\frac{1}{2}]$ and $q_1 \leq q_2 ... \leq q_N$. Let:
%
%\eqn {
%m \in [2,N] = \set{m:q_m > \frac{1}{m}}
%}
%
%$m$ is a measure of the number of infrequently occurring arms. If the problem is completely balanced $q_1...q_N = \frac{1}{2}$ then $m = 2$. If the problem is completely unbalanced, $q_1...q_N = 0$ then $m=N$.  Let $q_{ij} = \P{X_i = j}$ and $\mu_{ij} = P(Y|do(X_i = j))$.
%
%
%Suppose we select $I_t = 0$ for the first $h/2$ timesteps. This is at worst half the optimal. We use these observations to estimate the rewards for all the frequently occurring arms, $\set{do(X_i = 0) \; \forall i}$, and $\set{X_i = 1:i \geq m}$ 
%
%\eqn {
%\hat{\mu}_{ij} = \frac{\sum_{t=1}^{h/2}\ind{Y_t=1,X_{t,i}=j}}{\frac{h}{2}q_{ij}}
%}
%
%We then explicitly play each of the $m$ infrequently occurring arms $\frac{h}{2m}$ times and estimate their rewards as:
%
%\eqn {
%\hat{\mu}_{ij} = \frac{2m}{h}\sum_{t = 1}^{h/2m}\ind{Y_t = 1|X_i=j}
%}
%
%We then simply play the arm with the highest estimated reward for the remaining $T-h$ timesteps. We can show that we can select an $h$ (based on $T$, $N$ and $m$) that leads to worst case regret $R_T = \bigotilde{T^{2/3}m^{1/3}}$. This result can still be obtained in the case where $\boldsymbol{q}$ is unknown and must be learned during the exploration time. Comparing this to the standard bandit regret, $R_T = \bigo{\sqrt{TK}}$, we expect the causal algorithm to do better if $m < \frac{K^{3/2}}{\sqrt{T}}$. We can also show that the simple regret $R_s \in \bigotilde{\sqrt{m/T}}$. If $m << K$, this represents a significant improvement over standard bandit setting. \footnote{ $\bigotilde{.}$ and $\bigthetatilde{.}$ suppress polylogarithmic factors}
%
%
%In this algorithm, we have only used the side information provided by the $do()$ action about other actions. Since the $do()$ action fully reveals the value of alternate actions we could have incorporated this information via the graph feedback model \cite{Mannor2011}, where at each timestep the feedback graph $G_t$ is selected stochastically, dependent on $\boldsymbol{q}$, and revealed after an action has been chosen. The feedback graph is distinct from the causal graph. A link $A \rightarrow B$ in $G_t$ indicates that selecting the action $A$ reveals the reward for action $B$. For this specific problem, $G_t$ will always be a star graph with the action $do()$ connected to half the remaining actions. The Exp3-IX algorithm \cite{Kocak2014} was developed for the adversarial version of this problem and has regret $\bigo{\sqrt{\bar{\alpha}T}}$, where $\bar{\alpha}$ is the average independence number of $G_t$. In our case $\bar{\alpha} = \frac{N}{2}$ so we again obtain the regret of the standard bandit algorithm. The issue here is that a malicious adversary can select the same graph each time, such that the rewards for half the arms are never revealed by the informative action. This is equivalent to a, nominally, stochastic selection of feedback graph where $\boldsymbol{q} = \boldsymbol{0}$
%
%\cite{Lelarge2012} consider a stochastic version of the graph feedback problem, but with a fixed graph available to the algorithm before it must select an action. In addition, their algorithm is not optimal for all graph structures and fails, in particular, to provide improvements for star like graphs as in our case. \cite{Buccapatnam2014} improve the dependence of the algorithm on the graph structure but still assume the graph is fixed and available to the algorithm before the action is selected. 
%
%More generally, assuming causal structure creates more complex types of side information, such as that shown in equation \ref{eq:estimation_transfer}. In this case, selecting one action does not fully reveal an alternate action but provides some information towards an estimate. The quality of the estimate notably depends not only on the number of times that action was selected. For example, to get a good estimate for $X_1 = 1$ by intervening on $X_2$ requires us to sample both $X_2=0$ and $X_2=1$, in proportions dependent on $q_2$. This more complex side information does not fit within the graph feedback framework.
%
%Partial monitoring is a very general framework for for decoupling the feedback from the action and reward. It can be used to classify problems into one of four categories, trivial with no regret, easy with $R_T = \bigthetatilde{\sqrt{T}}$ , hard with $R_T = \bigtheta{T^{2/3}}$ and hopeless with $R_T = \bigomega{T}$ \cite{Bartok2014}. Partial monitoring algorithms yield results that are optimal with respect the the horizon $T$ but not other parameters, such as $K$, which is the key focus of incorporating causal structure. 
%
%
%
%\subsection{Open questions}
%
%There are many open questions still to be addressed. For the specific setting described above I need to derive lower bounds for the worst case regret so as to determine how close the current approach is to optimal. 
%
%It is clearly desirable to obtain an algorithm that could be generalized to an arbitrary causal graph. This will likely require the development of a bandit algorithm that can leverage more complex forms of side information like that in equation \ref{eq:estimation_transfer}, in addition to an efficient algorithm for eliminating actions that need not be considered at all.
%
%Another open class of problem to consider are cases where the causal structure is not (fully) known and also needs to be learnt within the online setting. 
%

\bibliographystyle{apalike}% Select the citation style e.g. ieeetr
\bibliography{../library}% write the directory to the .bib file
\end{document}

### 2016-06-08 causal-bandit/arXiv/general.tex
adds initial writup of contextual-bandit and identifiability problems

The second observation is that, in the parallel bandit setting, $m(\eta^*) \leq 2m(\boldsymbol{q})$. This is easy to see by letting $\eta_a = 1/2$ for $a = do()$ and $\eta_a = \ind{\P{X_i = j} \leq 1/m(\boldsymbol{q})} / 2m(\boldsymbol{q})$ for the actions corresponding to $do(X_i=j)$, and applying an argument like that for Proposition~\ref{pro:m-bound}.\ifsup 
The proof is in Section \ref{sec:m-equivelence}.
\else
The proof is in the supplementary materials.
\fi

>>>>

The second observation is that, in the parallel bandit setting, $m(\eta^*) \leq 2m(\boldsymbol{q})$. This is easy to see by letting $\eta_a = 1/2$ for $a = do()$ and $\eta_a = \ind{\P{X_i = j} \leq 1/m(\boldsymbol{q})} / 2m(\boldsymbol{q})$ for the actions corresponding to $do(X_i=j)$, and applying an argument like that for Proposition~\ref{pro:m-bound}. \ifsup 
The proof is in Section \ref{sec:m-equivelence}.
\else
The proof is in the supplementary materials.
\fi

### 2016-06-08 causal-bandit/arXiv/setup.tex
adds initial writup of contextual-bandit and identifiability problems

A learner for a casual bandit problem is given the casual model's graph $\mathcal{G}$ and a set of \emph{allowed interventions} $\mathcal{A}$.
One variable $Y \in \mathcal{X}$ is designated as the \emph{reward variable} and takes on values in $\{0, 1\}$.
We denote the expected reward for the intervention $a = do(\vec{X} = \vec{x})$ by $\mu_{a} := \E{Y | do(\vec{X} = \vec{x})}$ and 
the optimal expected reward by $\mu^* := \max_{a\in\actions} \mu_{a}$. 
The causal bandit game proceeds over $T$ rounds.
In round $t$, the learner \emph{intervenes} by choosing $a_t = do(\vec{X}_t = \vec{x}_t) \in \mathcal{A}$ based on previous observations. 
It then \emph{observes} sampled values for all non-intervened variables $\vec{X}^c_t$ drawn from $\P{\vec{X}^c_t | do(\vec{X}_t = \vec{x}_t)}$, 
including the \emph{reward} $Y_t \in \{0,1\}$. 
After $T$ observations the learner outputs an estimate of the optimal action $\hat a^*_T \in \actions$ based on its prior observations.

@@

We note that classical $K$-armed stochastic bandit problem can be recovered in our framework by considering a simple causal model with one edge connecting a single variable $X$ that can take on $K$ values to a reward variable $Y \in \set{0,1}$ where $\P{Y = 1|X} = r(X)$ for some arbitrary but unknown, real-valued function $r$. The set of allowed interventions in this case is $\mathcal{A} = \{ do(X = k) \colon k \in \{1, \ldots, K\}\}$. Conversely, any causal bandit problem can be reduced to a classical stochastic $|\mathcal{A}|$-armed bandit problem by treating each possible intervention as an independent arm and ignoring all sampled values for the observed variables except for the reward. Intuitively though, one would expect to perform better by making use of the extra structure and observations.

>>>>

A learner for a casual bandit problem is given the casual model's graph $\mathcal{G}$ and a set of \emph{allowed actions} $\mathcal{A}$.
One variable $Y \in \mathcal{X}$ is designated as the \emph{reward variable} and takes on values in $\{0, 1\}$.
We denote the expected reward for the action $a = do(\vec{X} = \vec{x})$ by $\mu_{a} := \E{Y | do(\vec{X} = \vec{x})}$ and 
the optimal expected reward by $\mu^* := \max_{a\in\actions} \mu_{a}$. 
The causal bandit game proceeds over $T$ rounds.
In round $t$, the learner \emph{intervenes} by choosing $a_t = do(\vec{X}_t = \vec{x}_t) \in \mathcal{A}$ based on previous observations. 
It then \emph{observes} sampled values for all non-intervened variables $\vec{X}^c_t$ drawn from $\P{\vec{X}^c_t | do(\vec{X}_t = \vec{x}_t)}$, 
including the \emph{reward} $Y_t \in \{0,1\}$. 
After $T$ observations the learner outputs an estimate of the optimal action $\hat a^*_T \in \actions$ based on its prior observations.

@@

We note that classical $K$-armed stochastic bandit problem can be recovered in our framework by considering a simple causal model with one edge connecting a single variable $X$ that can take on $K$ values to a reward variable $Y \in \set{0,1}$ where $\P{Y = 1|X} = r(X)$ for some arbitrary but unknown, real-valued function $r$. The set of allowed actions in this case is $\mathcal{A} = \{ do(X = k) \colon k \in \{1, \ldots, K\}\}$. Conversely, any causal bandit problem can be reduced to a classical stochastic $|\mathcal{A}|$-armed bandit problem by treating each possible intervention as an independent arm and ignoring all sampled values for the observed variables except for the reward. Intuitively though, one would expect to perform better by making use of the extra structure and observations.

### 2016-06-08 causal-bandit/arXiv/tech.tex
adds initial writup of contextual-bandit and identifiability problems

\begin{proof}
First note that $X_t, Y_t$ are sampled from $\operatorname{Q}$.
W define $Z_a(X_t) = Y_t R_a(X_t)\ind{R_a(X_t)\leq B_a}$ and abbreviate $Z_{at} = Z_a(X_t)$, $R_{at} = R_a(X_t)$ and $\Pn{a}{.} = \Pns{a}{.}$.
By definition we have $|Z_{at}| \leq B_a$ and 
\eq{
\Var_Q[Z_{at}] 
\leq \EE_Q[Z_{at}^2] 
\leq \EE_Q[R_{at}^2]
= \EEa[R_{at}]
= \EEa\left[\frac{\Pns{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right] 
\leq m(\eta)\,.
}
Checking the expectation we have
\eq{
\EE_Q[Z_{at}] 
= \EEa \left[Y \ind{R_{at} \leq B_a}\right] 
= \EEa Y - \EEa \left[Y\ind{R_{at} > B_a}\right] 
= \mu_a - \beta_a\,,
}
where 
\eq{
0 \leq \beta_a = \EEa[Y \ind{R_{at} > B_a}] \leq \Pns{a}{R_{at} > B_a}
}
is the negative bias. 
The bias may be bounded in terms of $m(\eta)$ via an application of Markov's inequality.
\eq{
\beta_a \leq \Pns{a}{R_{at} > B_a} \leq \frac{\EEa[R_{at}]}{B_a} \leq \frac{m(\eta)}{B_a}\,.
}
Let $\epsilon_a > 0$ be given by
\eq{
\epsilon = \sqrt{\frac{2m(\eta)}{T} \log\left(2T|\calA|\right)} + \frac{3B_a}{T} \log\left(2T|\calA|\right)\,.
}
Then by the union bound and Bernstein's inequality 
\eq{
\P{\text{exists } a \in \calA : \left|\hat \mu_a - \EE_Q[\hat \mu_a]\right| \geq \epsilon_a} 
\leq \sum_{a \in \calA} \P{\left|\hat \mu_a - \EE_Q[\hat \mu_a]\right| \geq \epsilon_a} \leq \frac{1}{T}\,.
}
Assuming this event does not occur and letting $a^* = \argmax_{a \in \calA} \mu_a$ we have
\eq{
\mu_I \geq \hat \mu_I - \epsilon_I  
\geq \hat \mu_{a^*} - \epsilon_I  
\geq \mu^* - \epsilon_{a^*} - \epsilon_I - \beta_{a^*}\,. 
}
By the definition of the truncation
we have
\eq{
\epsilon_a \leq \left(\sqrt{2} + 3\right)\sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)}
}
and
\eq{
\beta_a \leq \sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)}\,. 
}
Therefore for $C = \sqrt{2} + 4$ we have
\eq{
\P{\mu_I \geq \mu^* - C \sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)}} \leq \frac{1}{T}\,.
}
Therefore
\eq{
\mu^* - \EE[\mu_I] \leq C \sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)} + \frac{1}{T}
}
as required.
\end{proof}

>>>>

\begin{proof}
First note that $X_t, Y_t$ are sampled from $\operatorname{Q}$.
We define $Z_a(X_t) = Y_t R_a(X_t)\ind{R_a(X_t)\leq B_a}$ and abbreviate $Z_{at} = Z_a(X_t)$, $R_{at} = R_a(X_t)$ and $\Pn{a}{.} = \Pns{a}{.}$.
By definition we have $|Z_{at}| \leq B_a$ and 
\eq{
\Var_Q[Z_{at}] 
\leq \EE_Q[Z_{at}^2] 
\leq \EE_Q[R_{at}^2]
= \EEa[R_{at}]
= \EEa\left[\frac{\Pns{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right] 
\leq m(\eta)\,.
}
Checking the expectation we have
\eq{
\EE_Q[Z_{at}] 
= \EEa \left[Y \ind{R_{at} \leq B_a}\right] 
= \EEa Y - \EEa \left[Y\ind{R_{at} > B_a}\right] 
= \mu_a - \beta_a\,,
}
where 
\eq{
0 \leq \beta_a = \EEa[Y \ind{R_{at} > B_a}] \leq \Pns{a}{R_{at} > B_a}
}
is the negative bias. 
The bias may be bounded in terms of $m(\eta)$ via an application of Markov's inequality.
\eq{
\beta_a \leq \Pns{a}{R_{at} > B_a} \leq \frac{\EEa[R_{at}]}{B_a} \leq \frac{m(\eta)}{B_a}\,.
}
Let $\epsilon_a > 0$ be given by
\eq{
\epsilon_a = \sqrt{\frac{2m(\eta)}{T} \log\left(2T|\calA|\right)} + \frac{3B_a}{T} \log\left(2T|\calA|\right)\,.
}
Then by the union bound and Bernstein's inequality 
\eq{
\P{\text{exists } a \in \calA : \left|\hat \mu_a - \EE_Q[Z_{at}]\right| \geq \epsilon_a} 
\leq \sum_{a \in \calA} \P{\left|\hat \mu_a - \EE_Q[Z_{at}]\right| \geq \epsilon_a} \leq \frac{1}{T}\,.
}

Let $I = \hat{a}^*_T$ be the action selected by the algorithm, $a^* = \argmax_{a \in \calA} \mu_a$ be the true optimal action and recall that $\EE_Q[Z_{at}] = \mu_a - \beta_a$. Assuming the above event does not occur we have,

\eq{
\mu_I \geq \hat \mu_I - \epsilon_I  
\geq \hat \mu_{a^*} - \epsilon_I  
\geq \mu^* - \epsilon_{a^*} - \epsilon_I - \beta_{a^*}\,. 
}
By the definition of the truncation
we have
\eq{
\epsilon_a \leq \left(\sqrt{2} + 3\right)\sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)}
}
and
\eq{
\beta_a \leq \sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)}\,. 
}
Therefore for $C = \sqrt{2} + 4$ we have
\eq{
\P{\mu_I \geq \mu^* - C \sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)}} \leq \frac{1}{T}\,.
}
Therefore
\eq{
\mu^* - \EE[\mu_I] \leq C \sqrt{\frac{m(\eta)}{T} \log\left(2T|\calA|\right)} + \frac{1}{T}
}
as required.
\end{proof}
